---
title: "R Notebook"
output: html_notebook
---
For easier customization of the charts I have kept these functions using ggplot directly. Accessing the PCA analysis data this way could also be used to implement 3d graphs using 3 principal components, or using the other data calculated to produce a range of data visualisations

Perform pca
```{r}
prepare_pca_data <- function(snp_df){
  pca_analysis <- FactoMineR::PCA(snp_df[-(1:4)], graph = F) # assumes that first 4 columns are sample data so if not snps will be excluded
  pc1_pc2_df <- as.data.frame(pca_analysis$ind$coord[,1:2]) # selecting the first 2 principal components
  pc1_pc2_df <- cbind(snp_df[,1], pc1_pc2_df) %>% # adding back sampleID assuming they are present in col 1 of given df
    rename("snp_df[, 1]" = "sample_ID")
  pc1_pc2_df <- left_join(pc1_pc2_df, population_data, by = "sample_ID") # joining population data
  return(pc1_pc2_df)
}
```

Plot pca - standard pca plot with name selection, and colour by selection
```{r}
plot_pca <- function(pca_df, plot_name = "PCA plot", col_by = super_pop){
pca_plot <- pca_df %>% 
  ggplot() + geom_point(aes(x = Dim.1, y = Dim.2, color = {{col_by}})) + geom_hline(yintercept = 0, linetype = "longdash") + geom_vline(xintercept = 0, linetype = "longdash") + labs(title = paste(plot_name), y = "PC2", x = "PC1") # producing plot of principal components 1 & 2
print(pca_plot)
}
```

Quick pca - takes unprepared vcf data and calls on other functions to make everything from one line
```{r}
quick_pca <- function(vcf_data, plot_title = "Default_title"){
  plot_pca(prepare_pca_data(prepare_snp_data(vcf_data)),plot_title)
}
```

Multi pca analysis
```{r}
# plotting pca function - (Need to allow selection of colour by, currently colours by super_pop regardless)
# could add save function - Also may need to switch back from ggplot
multi_plot_pca <- function(pca_df, plot_name = "PCA plot"){
pca_plot <- pca_df %>% 
  ggplot() + geom_point(aes(x = Dim.1, y = Dim.2, color = super_pop)) + 
  geom_hline(yintercept = 0, linetype = "longdash") + geom_vline(xintercept = 0, linetype = "longdash") + 
  labs(title = paste(plot_name), y = "PC2", x = "PC1")

  print(pca_plot)
}


# running multi pca analysis - requires processed snps, vector to filter by (+ direction of what to filter by) and name is optional - Could filter by gender though there is only 3 combinations - M, F, MF
multi_pca <- function(df, selection_vector, filterby, naming = "PCA plot"){
for(i in 1:nrow(selection_vector)){ #looping for each row in vector table
  row_vector <- unlist(selection_vector[i,]) # Each row becomes filter criteria
  filtered_data <- df %>% 
    filter({{filterby}} %in% row_vector) # filtering snps by vector
  pcatst <- multi_plot_pca(prepare_pca_data(filtered_data), plot_name = paste0(i, ": ", naming)) # running pca and plotting for each criteria
  }
}
```